---
title: "Big Mountain Resort: Value Creation and Value Capture Strategy"
author: "John James john.james@nov8.ai"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: 
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    theme: flatly
    highlight: zenburn
    css: "style.css"    
---

```{r libraries, echo=FALSE, cache=FALSE, warning=FALSE, message=FALSE, results='hide'}
library(dplyr)
library(extrafont)
library(htmltools)
library(kableExtra)
library(kfigr)
library(knitr)
library(plotly)
library(reticulate)
library(rmarkdown)
library(tufte)
use_condaenv("sb")
matplotlib <- import("matplotlib")
matplotlib$use("Agg", force=TRUE)
```



```{r setup, echo=FALSE, cache=FALSE, warning=FALSE, message=FALSE}
figures <- '/figures/'
options(knitr.table.format = "html")
options(max.print=100, scipen=999, width = 800)
knitr::opts_chunk$set(echo=FALSE,
	             cache=FALSE,
               prompt=FALSE,
	             eval = TRUE,
               tidy=TRUE,
               fig.height = 8,
               fig.width = 20,
               comment=NA,
               message=FALSE,
               warning=FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_knit$set(width=100, figr.prefix = T, figr.link = T)
knitr::opts_chunk$set(fig.path = figures)
knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark=",")
})
```

```{python imports}
import numpy as np
import pandas as pd
import matplotlib
import matplotlib.pyplot as plt
import mlstudio
```

```{python syspath}
import inspect
import os
import sys
os.environ['QT_QPA_PLATFORM_PLUGIN_PATH'] = 'C:/Users/John/anaconda3/envs/gd383/Library/plugins/platforms'
# Add home directory to path
home = os.path.dirname(os.path.abspath(inspect.getfile(inspect.currentframe())))
sys.path.append(home)
```

# Big Mountain Current Situation         
Big Mountain Ski Resort offers spectacular views of Glacier National Park and Flathead National Forest, access to 105 trails, a vertical drop of 2353 feet, and over 3000 square feet of skiable terrain to some 350,000 visitors per year.  

The current asset base includes 2 T-bars, a magic carpet for novice skiers and 12 lifts, one of which was recently added to increase the distribution of visitors across the mountain resulting in a net annual operating cost increase of $1.54M.

Big Mountain's is currently evaluating how it creates and captures value vis-a-vis its investment, pricing, and operational strategies. The current pricing regime is to charge a premium over the average price of resorts in their market segment, but that approach is problematic in several respects. First, it provides no insight into what constitutes value from the perspective of the customer. Second, it doesn't reflect how and to what degree the resort delivers the value, and finally, the strategy doesn't maximize value capture for the company. Consequently, investment and cost-cutting strategies cannot be considered on the basis of valuable market and customer data.

Big Mountain Ski Resort has engaged outside council to 

2. propose a data-driven pricing strategy that reflects which aspects of Big Mountain's facilities and operations matter most to customers,    
1. illuminate the aspects of Big Mountain's facilities that have the greatest (and least) influence on price.    
2. propose a pricing strategy that maximizes revenue for the company and value for the customer, and           
3. provide data-driven recommendations vis-a-vis three shortlisted scenarios that the company is considering.

The analysis should help address questions such as:

1. How do we maximize price based upon the facilities currently in place?      
2. What investment strategies would create the most value for the company in the short and long run?            
3. How costs be cut without reducing value to the customer?     

This market and customer data will yield better investment and cost-cutting decisions, optimize value creation for the customer, and maximize value capture for Big Mountain! 

# Recommendation to Big Mountain: Add the Run!
Four scenarios are being considered by Big Mountain executives:

1. Permanently closing down up to 10 of the least used runs. This doesn't impact any other resort statistics.   
2. Increase the vertical drop by adding a run to a point 150 feet lower down but requiring the installation of an additional chair lift to bring skiers back up, without additional snow making coverage            
3. Same as number 2, but adding 2 acres of snow making cover             
4. Increase the longest run by 0.2 mile to boast 3.5 miles length, requiring an additional snow making coverage of 4 acres.

Of the four options being considered, adding the additional run and chair lift would support a ticket price increase of $1.99. With 350,000 visitors per year, this would correspond with an approximate \$3,5 million revenue increase per year. Adding the additional 2 acres of snow making had no incremental effect on price and increasing the longest run by .2 miles had no absolute impact on price.

# What Matters Most to Customers        

To assess the facilities that had the greatest influence on price, a feature correlation analysis and two machine learning models were developed. The features having the greatest influence in price were:

- vertical_drop    
- Snow Making_ac           
- total_chairs         
- fastQuads           
- Runs      
- LongestRun_mi       
- trams        
- SkiableTerrain_ac          

# What is Big Mountain's Price Position   
```{python comparisons, include=FALSE, code=readLines('./src/visualization/example.py')}
```
